---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# covid19mobility

<!-- badges: start -->
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![CRAN status](https://www.r-pkg.org/badges/version/covid19mobility)](https://CRAN.R-project.org/package=covid19mobility)
[![Travis build status](https://travis-ci.com/Covid19R/covid19mobility.svg?branch=master)](https://travis-ci.com/Covid19R/covid19mobility)
<!-- badges: end -->

The goal of covid19mobility is to make  mobility data from different sources available using the [Covid19R Project Data Format Standard](https://covid19r.github.io/documentation/data-format-standard.html). Currently, this package imports data from:
\

* The [Apple Mobility Trends Report](https://www.apple.com/covid19/mobility)

* The [Google Covid-19 Mobility Trends](https://www.google.com/covid19/mobility/)


## Installation
<!--
You can install the released version of covid19mobility from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("covid19mobility")
```
-->

You can install the current version of covid19mobility from github with:
```r
remotes::install_github("covid19r/covid19mobility")
```

## Data Format

The covid19mobility library follows the [Covid19R Project Data Format Standard](https://covid19r.github.io/documentation/data-format-standard.html), with some data sets holding extra data columns. To see what data is available, use `get_info_covid19mobility()`

```{r info}
library(covid19mobility)

get_info_covid19mobility() %>%
  dplyr::select(data_set_name, function_to_get_data, data_details) %>%
  knitr::kable()
```

The refresh methods bring in the different data sets. Currently available are:
* `refresh_covid19mobility_apple_country()` - [Apple Mobility Data](https://www.apple.com/covid19/mobility) at the country level.  
* `refresh_covid19mobility_subregion()` - [Apple Mobility Data](https://www.apple.com/covid19/mobility) at the state/subregion level.  
* `refresh_covid19mobility_apple_city()` - [Apple Mobility Data](https://www.apple.com/covid19/mobility) at the city level. Contains some lat/longs for some cities.
* `refresh_covid19mobility_google_country()` - [Google Mobility Data](https://www.google.com/covid19/mobility/) at the country level.  
* `refresh_covid19mobility_google_subregions()` - [Google Mobility Data](https://www.google.com/covid19/mobility/) at the state/subregion level.  
* `refresh_covid19mobility_google_us_counties()` - [Google Mobility Data](https://www.google.com/covid19/mobility/) at the US county level with FIPS codes.

For example

```{r county, message=FALSE, warning=FALSE}
refresh_covid19mobility_google_us_counties() %>%
  head()
```

## Data Use and Licensing

Please see the relevant information about each data set and their licenses if you plan on using these data in any published works.
